---
editor: 
  markdown: 
    wrap: 72
---

# Social network models

Our jobs, opportunities, political opinions, and many other personal
attributes are constrained by our social networks. A *social network* is
a mathematical, theoretical representation of relationships between
individual entities including people, or between companies,
institutions, governments, or other organizations. Relationships could
have many forms: friendship, economic exchange, or instruction/learning
are all forms of relationships that could be represented as a social
network. The word *network* is just the long form of the word *net*, as
in a fishing, mosquito, or storage net [@Smaldino2023, Ch. 9]. In
Portuguese, for example, the phrase *social network* is
translated *rede social* (*red social* in Spanish), where *rede* (*red*) 
translates to *net* when not used to describe social net(works).

## Basic network model concepts

In social networks, each individual (or individual entity) is
represented by a *vertex* or *node*. Vertices are essentially containers
to organize an individual's attributes including a name or other unique
identifier. For example, when representing a person in a friendship
network, a vertex could have attributes *name*, *age*, *weight*,
*height*, *socio-economic status*, *nationality*, *ethnicity*, *gender*,
*political party*, whether they *like roller coasters*, etc. Vertices
are drawn as circles, or another shape, whose size or color could be
related to one of the attributes, e.g., circle size could represent
income and circle color could represent political party.

![Arrows represent the flow of information, money, influence, etc., in asymmetric social relationships (A). 
Lines with no arrows, or two arrows between agents pointing one in each
direction, represent symmetric relationships
(B). Edges can represent different relationships; e.g., 
$i$ teaches karate to $j$ (pink arrow), and in return $j$ gives $i$ hand-made furniture 
(orange arrow) (C). 
Nodes can represent different entity types, e.g.,
$i$ and $j$ work for and are paid by company $k$
(D).](images/clipboard-3994932320.png){style="float: left; margin: 5px;"
width="250" #fig-vertex-edge-example}

Relationships in social networks between individuals are represented by
*edges*, drawn as lines between vertices [@fig-vertex-edge-example]. *Asymmetric relationships*, such as teacher-learner, seller-buyer, etc., involve the flow of something, like information or products/services in these examples. In this case the lines between vertices have arrows. Symmetric relationships, such as healthy friendships and geographic proximity, are drawn as a single line without arrows.
We can say there exists a *social connection* if $i$ and $j$ have some sort of
social, economic, or other relationship with one another, or we can say they
are *connected*. When we create social network models we will always specify 
what being connected means. When it comes to the diffusion of adaptations, the


### Simple example

INSERT SIMPLE 4-PERSON NEIGHBORHOOD HERE

## The importance of social networks

- Development and diffusion of innovations and adaptations, 
    - @Derex2016, @Moser2023
    - core-periphery ideas:
        - @ 
    - Networks in epidemiology: 
        - sexual networks for STD prediction and prevention  
        - Employment and travel patterns for respiratory infections
- Political polarization and social influence
    - @Turner2018
    - @Jackson2008, Ch. 8, starting on p. 223


## Working with network models



### Create a network from scratch


```{r}
#| output: false
library(igraph)
library(ggnetwork) 
```
1. Create an empty graph
1. Add nodes with attributes 
1. Add edges
1. Visualize the network using [`ggnetwork`](https://briatte.github.io/ggnetwork/index.html) for a ggplot interface for graph viz

This approach may be useful if you want to create a stylized network but aren't
aware of any existing data.
You could specify that there exist some number of "leader" nodes who have more out-edges than in-edges, for example, representing advice, instructions, or other information that more frequently flows from leaders to others. 




### Load network data from a library

This example shows how to load network data from an R library. Specifically,
we will load the Florentine intermarriage network data used as a motivating
example for the introductory chapter to @Jackson2008. First, we load a
few more R packages we'll need for this.


```{r}
#| output: false
library(netrankr)  # To load Florentine marriage network.
library(tibble)    # To make ggnetwork-gen'd list into a tibble.
library(dplyr)     # To easily rename tibble columns.
```


```{r}
#| label: fig-florentine-families
#| caption: 
#| fig-cap-location: margin
#| fig-height: 3

# The Florentine network data contains one disconnected family; remove it here.
florentine_m <- delete_vertices(florentine_m, which(degree(florentine_m) == 0))

# Convert to ggnetwork format for plotting.
flornet <- 
  tibble(ggnetwork(florentine_m)) %>% rename(Name = name, Wealth = wealth)

# The x, y, xend, and yend variables are assigned in ggnetwork call.
ggplot(flornet, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey", ) +
  geom_nodes(aes(size = Wealth), color = "#8AB8A7") + 
  geom_nodelabel_repel(aes(label = Name), color = "#0099b7", 
                       box.padding = unit(0.5, "lines")) +
  theme_blank()
```

### Network measures

#### Distance

#### Transitivity

#### Betweenness 

#### Exposure probability


## Regular networks


## Random networks 


## Homophily

The final type of network construction process is somewhat more
empirically-motivated than others: this is the *homophily*-structured network.
Homophily is the name given to a network measure of how likely individuals
are to have a social connection

- Humans evolved to form groups [@Cikara2014;@Cikara2017]
  - @EhretConstantino2022
- @Centola2010, @Centola2011 to introduce the concept and importance of homophily in a series of interventions against obesity. Obese individuals were more successful adopting practices that limited obesity when homophily was greater and they interacted more frequently with other obese people 
- Two types of homophily:
  - *Choice homophily*: Personal preferences for interacting with similar others, not necessarily conscious (see implicit bias stuff)
  - *Structural* or *induced* homophily: A history of racial segregation, one's career, and similar factors affect how frequently one interacts with in-group members versus out-group members
- In Centola's theory, homophily is a global attribute assumed equal for all individuals. In fact, it's specified by the experimental design.
- The more general case is asymmetric homophily, where homophily is set based
on group membership [@Morris1991;@Lee2019;@Turner2023]
- The most general case, rarely modeled, is 

### Construction of homophilic networks

```{r}

```
